---
title: "getMetlinVignette"
author: "William Kumler"
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("getMetlinFunctions.R")
```

### Intro

Metlin's got a ton of useful data, but it's trapped behind a weird interactive
Javascript thing in their website. The functions here allow the retrieval of
arbitrary amounts of data from the website in a scripted fashion.

NOTE: Metlin has some rate-limitations in place that, if exceeded, will block an IP for about a day. I triggered this while trying to pull down the data for 100 compounds within a minute, and haven't played with it beyond that. I'm not sure how generous or aggressive the limits actually are.

### Get data for a given compound name or mass

When called, **getMetlinMz** and **getMetlinName** will print out some diagnostic
information about the number of compounds found and general structural info. The
object *returned* by the function is a data frame with columns for Metlin's internal
compound ID, the exact mass of the compound, and the compound name, formula,
CAS and KEGG numbers. The final column tells you whether they have MS/MS data for
the given compound.

**getMetlinName** tends to be a little overeager, and may return compounds that
are only vaguely related to your search. This is a function of Metlin's name
search algorithm, not this code, and you'd get the same results via the Advanced
Search function on the Metlin website.

```{r}
mass_data <- getMetlinMz(117.078979)
str(mass_data)
head(mass_data)

name_data <- getMetlinName("Betaine")
str(name_data)
head(name_data)
```

### Get MS/MS data for a given compound

After a preliminary search has been established, compounds with "experimental"
in the MSMS column can be searched again for the MS/MS data specifically. 
**getMetlinMS2** accepts the unique Metlin ID for a compound and returns a
data frame with all MS/MS data in long format. The columns are "polarity" (either
"+" or "-", depending on the polarity the data was collected in), "adduct" (
usually "[M+H]" or "[M-H]"), "voltage" (the voltage of the collision cell - 
higher voltages typically produce smaller fragments), and finally each fragment's
mass and intensity as "frag_mass" and "frag_int", respectively. Intensity is
normalized to the highest peak, and therefore scaled from 100.

```{r}
#Preliminary search for a compound
name_data <- getMetlinName("Betaine")

#Access compound ID for the actual compound we'd like to get MS/MS data for
betaine_data <- name_data[3,]
betaine_id <- as.numeric(as.character(betaine_data$cmpd_id))

#Request ms2 data
betaine_ms2 <- getMetlinMS2(betaine_id)
str(betaine_ms2)
head(betaine_ms2)
```

### Plot the data (base R)

```{r}
# Base R plotting
# Split the data frame into each voltage and look at only one polarity
split_volt_pos_ms2 <- betaine_ms2 %>%
  subset(polarity=="+") %>%
  split(.$voltage)
# Arrange the output nicely
layout(matrix(c(1, rep(2:(length(split_volt_pos_ms2)+1), each=2),  1), ncol = 1))
par(mar=c(0.1, 4.1, 0.1, 0.1))
plot.new()
# Aaaaaand plot each voltage
text(x = 0.5, y=1, labels = betaine_data$cmpd_name, cex=3)
for(i in split_volt_pos_ms2){
  plot(i$frag_mass, i$frag_int, xlab = "", 
       ylab=paste("Voltage", unique(i$voltage)),
       xlim=c(0, max(betaine_ms2$frag_mass)), 
       xaxt="n", yaxt="n", type="n", ylim=c(0, 120))
  segments(x0 = i$frag_mass, x1 = i$frag_mass,
           y0 = 0, y1 = i$frag_int)
  axis(side = 2, at = c(0, 50, 100), labels = c(0, 50, 100))
}
axis(side = 1)
```

### Plot the data (ggplot method)

```{r}
library(ggplot2)
betaine_ms2 %>%
  filter(polarity=="+") %>%
    #filter(voltage==20) %>%
    ggplot(label=frag_mass) +
    geom_segment(aes(yend=0, x=frag_mass, y=frag_int, xend=frag_mass)) +
    #geom_hline(yintercept=0) +
    facet_wrap(~voltage, ncol = 1) +
    theme_bw() +
    xlim(0, max(betaine_ms2$frag_mass))
```

The ggplots can also be saved and rendered interactively with **ggplotly**, to
mimic the interactive method available on the Metlin website.