---
title: "getMetlinVignette"
author: "wkumler"
date: "September 30, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Metlin's got a ton of useful data, but it's trapped behind a weird interactive
Javascript thing in their website. The functions here allow the retrieval of
arbitrary amounts of data from the website in a scripted fashion.

Two notes - one, I'm pretty sure this would not make Metlin happy. I'm not sure
it's technically *illegal* but I'd rather not have to figure that out. Two,
Metlin has some rate-limitations in place that, if exceeded, will block an IP
for about a day. I triggered this while trying to pull down the data for
20 amino acids, 5 times, within a minute. So I'm not sure how generous or 
aggressive they actually are.

```{r cars}
library(getMetlin)
library(dplyr)

# Start by searching for Metlin records by name or mass
head(getMetlinMz(117.078979))
head(getMetlinName("Betaine"))

# Then, access the MS/MS data for a record that has them
sample_ms2_cmpd <- getMetlinMz(117.078979) %>% 
  filter(MSMS=="experimental") %>% 
  slice(1)

metlin_id <- as.numeric(as.character(sample_ms2_cmpd$cmpd_id))

sample_ms2 <- getMetlinMS2(metlin_id)
head(sample_ms2)

# Plot it
split_volt_pos_ms2 <- sample_ms2 %>%
  subset(polarity=="+") %>%
  split(.$voltage)
layout(matrix(c(1, rep(2:(length(split_volt_pos_ms2)+1), each=2),  1), ncol = 1))
par(mar=c(0.1, 4.1, 0.1, 0.1))
plot.new()
text(x = 0.5, y=1, labels = sample_ms2_cmpd$cmpd_name, cex=3)
for(i in split_volt_pos_ms2){
  plot(i$frag_mass, i$frag_int, xlab = "", 
       ylab=paste("Voltage", unique(i$voltage)),
       xlim=c(0, max(sample_ms2$frag_mass)), 
       xaxt="n", yaxt="n", type="n", ylim=c(0, 120))
  segments(x0 = i$frag_mass, x1 = i$frag_mass,
           y0 = 0, y1 = i$frag_int)
  axis(side = 2, at = c(0, 50, 100), labels = c(0, 50, 100))
}
axis(side = 1)

library(ggplot2)
gp <- sample_ms2 %>%
  filter(polarity=="+") %>%
    #filter(voltage==20) %>%
    ggplot(label=frag_mass) +
    geom_segment(aes(yend=0, x=frag_mass, y=frag_int, xend=frag_mass)) +
    #geom_hline(yintercept=0) +
    facet_wrap(~voltage, ncol = 1) +
    theme_bw() +
    xlim(0, max(sample_ms2$frag_mass))
gp

library(plotly)
ggplotly(gp, tooltip = c("y", "x"))

```
